How is experience transformed into memory?

### 2018-08-23 paper/main.tex
working on intro

\begin{abstract}
The human memory system is adept at cataloging the rich dynamics of ongoing experience. A defining feature of our everyday experiences is that, despite the rich complexities of real-world experiences, they unfold in a structured and often predictable manner. This structure gives each of our experiences (and thus our memories for those experiences) a unique `shape' with respect to how it unfolds in the relevant representational space.  However, temporal structure is typically absent (or severely impoverished) in traditional trial-based  experiments. As a consequence, our understanding of how the temporal structure (i.e., the shape) of our experiences influences memorability and memory organization is limited. In this study, we investigate how people verbally recount a video by characterizing and relating the thematic dynamics, or ``trajectories,'' of the video to participants' later recounting of the narrative. Using a novel approach that leverages topic modeling and Hidden Markov Models, we found that, despite substantial variability in the quantity of information recalled across participants as well as the particular words used to describe the experience, the overall shape of the stimulus was recapitulated by nearly every participant. These findings provide a window into which aspects of naturalistic experiences must be preserved, and which might be more flexible, in considering whether and how those experiences are remembered.
\end{abstract}

@@

What does it mean to \textit{remember} something? In traditional episodic memory experiments \citep[e.g., list-learning or trial-based experiments;][]{Murd62a, Kaha96}, remembering is often cast as a binary operation: either an item is recalled or it isn't. More nuanced studies might incorporate self-reported confidence measures as a proxy for memory strength, or ask participants to discriminate between ``recollecting'' an experience or a feeling of ``familiarity'' \citep{Yone02}. However, characterizing and evaluating memory in more realistic contexts (e.g., telling a story to a friend about a movie you recently saw) is much more nuanced. Real-world recall is continuous, rather than binary.  The specific words used to describe an experience have very little bearing on whether the experience is considered to have been ``remembered.''  Further, one might remember the gist of an experience but forget (or neglect to recount) particular details.  Or different people who share an experience might recount the experience with a similar level of detail, but the specific details that were remembered might vary across people.  Which aspects of those recollections should be considered fundamental, and which are extraneous to the main story?

>>>>

\begin{abstract}
The human memory system is adept at cataloging the rich dynamics of ongoing experience. A defining feature of our everyday experiences is that, despite the rich complexities of real-world experiences, they unfold in a structured and often predictable manner. This structure gives each of our experiences (and thus our memories for those experiences) a unique `shape' with respect to how it unfolds in the relevant representational space.  However, temporal structure is typically absent (or severely impoverished) in traditional trial-based  experiments. As a consequence, our understanding of how the temporal structure (i.e., the shape) of our experiences influences memorability and memory organization is limited. In this study, we investigate how people verbally recount a video by characterizing and relating the thematic dynamics, or ``trajectories,'' of the video to participants' later recounting of the narrative. Using a novel approach that leverages topic modeling and Hidden Markov Models, we found that, despite substantial variability in the quantity of information recalled across participants as well as the particular words used to describe the experience, the overall shape of the stimulus was recapitulated by nearly every participant. Our findings suggest that assessing the degree of match between the shape of an experience and the subsequent retelling of that experience provides a much more nuanced view (compared with traditional methods, such as comparing the proportion of remembered details) into what it means to remember. These findings provide a window into which aspects of naturalistic experiences must be preserved, and which might be more flexible, in considering whether and how those experiences are remembered.
\end{abstract}

@@

What does it mean to \textit{remember} something? In traditional episodic memory experiments \citep[e.g., list-learning or trial-based experiments;][]{Murd62a, Kaha96}, remembering is often cast as a binary operation: either a studied item is recalled or it isn't. More nuanced studies might incorporate self-reported confidence measures as a proxy for memory strength, or ask participants to discriminate between ``recollecting'' an experience or a general feeling of ``familiarity'' \citep{Yone02}. However, characterizing and evaluating memory in more realistic contexts (e.g., recounting a recent experience to a friend) concerns a much different set of criteria.  First, eal-world recall is continuous, rather than binary.  Second, the specific words used to describe an experience have very little bearing on whether the experience is considered to have been ``remembered.''  Third, one might remember the gist of an experience but forget (or neglect to recount) particular details.  Or different people who share an experience might recount the experience with a similar level of detail, but the specific details that were remembered might vary across people.  Which aspects of those recollections should be considered fundamental to determining whether or how reliably an experience has been remembered?

%JRM STOPPED INPUTTING EDITS HERE
